---
title: "ESJ Map"
author: "Brian"
date: "2/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cluster)
library(sf)
library(viridis)
library(leaflet)
library(mapview)
library(tmap)
```

```{r}
kc_demographics <- read_rds("kc_demographics.rds")
#glimpse(kc_demographics, width = 100)
head(kc_demographics)
```

```{r}
set.seed(23994)
vars <- c('med_income', 'p_e_less', 'percent_poc')#variables needed
equity_matrix <- cluster::daisy(kc_demographics[,vars])
res.pam <- cluster::pam(x=equity_matrix,k = 5,cluster.only = F)
#test
kc_demographics$cluster <- as.double(res.pam$clustering)
#kc_demographics$cluster <- as.factor(res.pam$clustering)
#kc_demographics$cluster=factor(dataCSV$cluster,
#                           levels=c(1,3,2),
#                           labels=c("low","mid","best"), 
#                           ordered=T)
head(kc_demographics)
```

# Notes form 2/25/22 meeting with Prof. Magallanes

## Whats going on when i cluster?

relevant when we have several variables and we want to make a summary

counter to the arithmetic mechanism for summary

it is based on the distance between each tract based on a distance matrix

when it is all numeric, then we are doing euclidian distance

other is gauer distance (spelling) - for when there is data of multi types

powerful if you have many variables

## When does clustering happen

PAM - is the algorithm that makes the cluster. It starts with pairs that are maximally distant, then get other centoids to the max distances. can also be used for categorical data more resistant to outliers

others include k means and others (family is the partitioning technique)

## what are the pitfalls

good to standardize the variables before you cluster

a poor choice of variables

we need good theory behind caribales to define the clusters

fronteers, things on the edges of groups

**these are the negative silloets**

## Checking the goodness of fit

silloets

\^\^ but it may not be good for longitudinal analysis

\#\#What are "natural groups"

it may be "Jenks Natural Breaks Classification"

this requires some arthmetic manipulation to get the thee varibales to make one variable that can be brokin in to categories

\#\#Where can I learn more?

\#\#how to get this to knit?

Exploring the data

c_demographics %\>% glimpse(width = 100)

```{r}
kc_demographics %>% glimpse(width = 100)
table(kc_demographics$cluster,
                exclude = 'nothing')
summary(kc_demographics$cluster)
summary_income <- aggregate(data=kc_demographics, med_income~cluster, median)
summary_english <- aggregate(data=kc_demographics, p_e_less~cluster, mean)
summary_poc <- aggregate(data=kc_demographics, percent_poc~cluster, mean)
esj_summary <- summary_income %>% inner_join(summary_english) %>% inner_join(summary_poc)
esj_summary[order(esj_summary$percent_poc),]
```

```{r}
kc_demographics$cluster <- factor(kc_demographics$cluster,
                                  levels=c(1,5,2,3,4),
                                  labels=c("low income diverse",
                                           "low income middle diverse",
                                           "middle income middle diverse", 
                                           "high income low diversity", 
                                           "middle income low diversity"),
                                  ordered=T)
kc_demographics
```

```{r}
ESJ_map <- kc_demographics %>% mutate(esj_score = as.integer(cluster)) %>%
  st_sf() %>%
  ggplot(
    aes(fill = esj_score)) +
  geom_sf() +
  scale_fill_viridis()
ESJ_map
```
```{r}
roadsproj=read.csv("Geographic Data Roads Projects.csv")
```

```{r}
roadssub=roadsproj[complete.cases(roadsproj$Latitude),]
```
```{r}
RoadsMap = ESJ_map + geom_sf(data = roadssub, aes(x = Latitude, y = Longitude), color=NA)
```





